---
title: "Final-Project-2: Classification Package Vignette"
author: "Shahab Musavi, John Grace, Nabin Neupane"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Classification Package Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Classification)

```


## Introduction
The classification package provides tools for logistic regression using numerical optimization, computing bootstrap confidence intervals for model coefficients, and evaluating model performance metrics.
This vignette demonstrates these functionalities using the adult dataset.

## Preprocessing the Data

### Dataset Overview

First, let let us take a look at the first few rows of the adult data set. The age values are re-scaled using the max-min method to make them fall into the range of [0,1] to match other binary values. 

- **1** indicates income > 50K
- **0** indicates income ≤ 50K


```{r}
# Load the dataset
adult <- read.csv("adult.csv")
head(adult)
```

### Feature Selection and Transformation

The adult dataset contains categorical variables, so we first preprocess it to convert categorical data into numerical values. Also, to reduce the run time, we just consider *age* and *education* as the predictors.

```{r}
# Select the 'age' and 'education' columns as features and 'income' as the outcome
features <- adult[, c("age", "education")]
outcome <- adult$income  # Assuming 'income' is the binary response variable

# One-hot encode the 'education' column (categorical variable)
education_transformed <- model.matrix(~ education - 1, data = features)

# Combine the 'age' column with the one-hot encoded 'education' columns
X <- cbind(1, features$age, education_transformed)  # Add an intercept column

# Prepare y (response variable)
y <- outcome  # Keep the outcome variable as-is

# Check dimensions
dim(X)  # Verify rows (observations) and columns (intercept + predictors)
length(y)  # Verify length matches rows in X


```

## Logistic Regression

We fit a logistic regression model using the ```logistic_regression()``` function. Here, we only consider the first 1000 rows of X and y to reduce the run time.

```{r}

# Subset the first 1000 rows of X and y
X_subset <- X[1:1000, ]
y_subset <- y[1:1000]

# Fit logistic regression on the subset
beta <- logistic_regression(X_subset, y_subset)
print(beta)

```

## Bootstrap Confidence Intervals
We compute 95% confidence intervals for the coefficients using the ```bootstrap_ci()``` function.

```{r}
ci <- bootstrap_ci(X_subset, y_subset, n_boot = 50, alpha = 0.05)
print(ci)
```

## Model Evaluation
We evaluate the performance of the logistic regression model using the ```evaluate_model()``` function.

```{r}
# Predicted probabilities
y_pred <- 1 / (1 + exp(-X %*% beta))

# Evaluate model
metrics <- evaluate_model(y, y_pred, cutoff = 0.5)
print(metrics)

```

## Summary 

This vignette demonstrated:
 
- Preprocessing: Data preparation and transformation.
- Modeling: Logistic regression for binary classification.
- Inference: Bootstrap confidence intervals for coefficients.
- Evaluation: Metrics to assess model performance.

Find more information here: https://au-r-programming.github.io/Final-Project-2/

## References

1- https://medium.com/@curryrowan/simplified-logistic-regression-classification-with-categorical-variables-in-python-1ce50c4b137

2- https://smac-group.github.io/irg/articles/vignette.html

3- https://stackoverflow.com/questions/72179150/error-in-read-dcf-line-starting-is-malformed

4- https://chatgpt.com/share/67532e2d-2d1c-8011-b0a6-0c85340e9182




